Attended

### 2021-10-13 draft_paper.Rmd
added time on scene stuff

# read in forecasts
# path to UoM server different for everyone
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")
domestic_by_grade_volume_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_volume_forecasts.Rds")
domestic_by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/domestic_by_grade_rt_forecasts.Rds")
domestic_by_grade_attended_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_attended_forecasts.Rds")

@@

  # conditionally label y axis
  y_axis_label <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast", 
                                           "domestic_by_grade_rt_forecast") ~ "Median response time (in minutes)", 
                         df_name %in% c("call_forecasts", 
                                           "domestic_by_grade_volume_forecasts") ~ "Weekly count of calls for service", 
                         df_name %in% c("by_type_attended_forecast",
                                        "by_grade_attended_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "Weekly % of calls attended")

  # conditionally find actual value column
  actual_col_name <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast",
                                           "domestic_by_grade_rt_forecast") ~ "actual_rt", 
                         df_name %in% c("call_forecasts", 
                                           "domestic_by_grade_volume_forecasts") ~ "actual_calls", 
                         df_name %in% c("by_type_attended_forecast",
                                        "by_grade_attended_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "actual_attended")

  forecasts %>% 
    filter(eval(as.symbol(variable)) == value) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, eval(as.symbol(actual_col_name)))) +
    geom_point(aes(incident_week, eval(as.symbol(actual_col_name)), fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21")))+
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = value,
      subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = y_axis_label,
      fill = "Actual observed value significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

@@

The contribution of this paper is three fold. First, we contribute to the growing literature which aims to better understand demand on police time, and how the pandemic and associated lockdowns influenced this. Second, we disaggregate calls from areas with different socio-demographic characteristic, to understand not only how police demand changed, but the variations in this change between different groups in society. Finally, we showcase temporal forecasting as an appropriate methodological approach to compare what we might have expected to happen in 2020/2021 in the absence of the pandemic, versus the observed data.

@@

# Total calls fig

@@

# % attended fig and response time fig

@@

@@

@@

Attend more stuff where less time at scene - eg night time economy calls takes people off scene for longer? *check time on scene*

With fewer calls for *list everything that had fewer than pre-covid* it is likely that officer workload changed (i.e. lowered). This meant quicker response times for calls. Figure X shows median response times by grade. Response times decreased for grade 1, grade 2, and grade 3 calls, the most drastic decrease in the latter category. This suggests that the freed up resources allowed faster responding to routine response calls. There was no difference in response time to those grade 4 calls (these require attendance or other resolution within 48 hours) and grade 5 calls are not attended.

>>>>

# read in forecasts
# path to UoM server different for everyone
all_calls <- read_rds("/Volumes/n8_covid/all_calls_forecast.Rds")
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")
attended_by_grade_forecast <- read_rds("/Volumes/n8_covid/attended_by_grade_forecasts.Rds")
domestic_by_grade_volume_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_volume_forecasts.Rds")
domestic_by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/domestic_by_grade_rt_forecasts.Rds")
domestic_by_grade_attended_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_attended_forecasts.Rds")
by_grade_tos_forecasts <- read_rds("/Volumes/n8_covid/by_grade_tos_forecasts.Rds")

@@

  # conditionally label y axis
  y_axis_label <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast", 
                                           "domestic_by_grade_rt_forecast") ~ "Median response time (in minutes)", 
                         df_name %in% c("call_forecasts", 
                                        "domestic_by_grade_volume_forecasts") ~ "Weekly count of calls for service", 
                         df_name %in% c("by_type_attended_forecast",
                                        "attended_by_grade_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "Weekly % of calls attended", 
                         df_name %in% c("by_grade_tos_forecasts") ~ "Median time on scene (in minutes)")

  # conditionally find actual value column
  actual_col_name <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast",
                                           "domestic_by_grade_rt_forecast") ~ "actual_rt", 
                         df_name %in% c("call_forecasts",
                                        "domestic_by_grade_volume_forecasts") ~ "actual_calls", 
                         df_name %in% c("by_type_attended_forecast",
                                        "attended_by_grade_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "actual_attended", 
                         df_name %in% c("by_grade_tos_forecasts") ~ "actual_tos")

  forecasts %>% 
    filter(eval(as.symbol(variable)) %in% value) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, eval(as.symbol(actual_col_name)))) +
    geom_point(aes(incident_week, eval(as.symbol(actual_col_name)), fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21")))+
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = value,
      subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = y_axis_label,
      fill = "Actual observed value significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

@@

The contribution of this paper is three fold. First, we contribute to the growing literature which aims to better understand demand on police time, and how the pandemic and associated lockdowns influenced this. Second, we disaggregate calls from areas with different socio-demographic characteristic, to understand not only how police demand changed, but the variations in this change between different groups in society. Finally, we showcase temporal forecasting as an appropriate methodological approach to compare what we might have expected to happen in 2020 in the absence of the pandemic, versus the observed data.

@@

```{r totalcallsfig}

# Total calls fig
all_calls %>% 
 mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
  ggplot() +
  # Forecast
  geom_ribbon(
    aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
    na.rm = TRUE,
    alpha = 0.5, 
    fill = "grey80"
  ) +
  geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
  # Dates of interest
  geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
  geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
  # Actual calls
  geom_line(aes(incident_week, actual_calls)) +
  geom_point(aes(incident_week, actual_calls, fill = sig), shape = 21) +
  scale_x_date(date_labels = "%e %b\n%Y", 
               limits = as.Date(c("2020-01-06", "2020-12-21"))) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
  scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
  labs(
    title = "Change in all calls for service during 2020 compared to pre-pandemic forecast",
    subtitle = str_wrap(
      str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
      80
    ),
    caption = "Forecast calculated using data up to 31 January 2020",
    x = NULL,
    y = "Weekly number of calls",
    fill = "actual calls significantly different from forecast"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold")) + 
    xlim(c(min(all_calls$incident_week), ymd("2020-12-14")))

@@

```{r allattendedfig}

# % attended fig and response time fig
all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2020 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

```{r allresptimefig}

all_rt_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_rt)) +
    geom_point(aes(incident_week, actual_rt, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Median response time for all calls for service during 2020 compared to pre-pandemic forecast",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "Median response time in minutes",
      fill = "Response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

@@

With fewer calls it is likely that officer workload changed (i.e. lowered). This means that we see police responding to a higher proportion of calls than we forecast with the absence of lockdown. Indeed, we tend to see this higher proportion of calls being attended for grade 1 (Emergency Response: Attendance within 15 minutes of call receipt), grade 2 (Priority Response: Attendance within 1 hour of call receipt), grade 3 calls (Routine Response: Attendance within 4 hours), and even in grade 4 calls (Scheduled Response: Attendance or other resolution within 48 hours) (grade 5 calls are not attended, these refer to Telephone Resolution: First-time telephone resolution of a call).

```{r bygradeattendedperc}

forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 1")
forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 2")
forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 3")
forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 4")

As well as being able to respond to more calls, 2020 also saw quicker response times for calls. Figure X shows median response times by grade. Response times decreased for grade 1, grade 2, and grade 3 calls, the most drastic decrease in the latter category. This suggests that the freed up resources allowed faster responding to routine response calls. There was no difference in response time to those grade 4 calls (these require attendance or other resolution within 48 hours) and grade 5 calls are not attended.

```{r bygraderesptime}

Another possible explanation for the increased time besides the decrease in call volume is that the calls which did come in required less time on scene.

```{r bygradetosfigs}

forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 1")
forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 2")
forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 3")
forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 4")

So not only were fewer calls coming in overall, which required less time to be spent on scene (??) but also the route to them was not so much blocked with traffic.

@@

The quiet streets allowed not only for faster response times, but also for an ease of spotting people out and about doing things they should not be doing.

@@

### 2021-10-13 output/attending_or_not.Rmd
added time on scene stuff

all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

>>>>

all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2020 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))
